---
title: 'P8105: Homework 3'
author: "Derek Lamb"
date: "`r Sys.Date()`"
output: github_document
---

I will load all necessary packages here. I also am including the code that Dr. Goldsmith showed in class, to set figure color and size preferences. I have modified it slightly: I prefer `theme_bw()` over `theme_minimal()` and I may change the color scheme from the default `viridis`.
```{r load packages, message = FALSE}
library(tidyverse)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

## These are options that i'm going to treat as default for now
## I don't love viridis, but don't have a better option yet
theme_set(theme_bw() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
```{r instacart data import}
# Load data from the provided package
library(p8105.datasets)
data("instacart")
```

### Data set size, user behavior
This data set is massive, containing `r nrow(instacart)` observations of `r ncol(instacart)` variables. Each row is a specific item from an order, in total `r instacart |> pull(order_id) |> unique() |> length()` orders are included from `r instacart |> pull(user_id) |> unique() |> length()` unique users. There appear to be no repeat orders from users in this data set, although `r instacart |> pull(reordered) |> sum()` items are items that were previously ordered by that user. This accounts for `r (instacart |> pull(reordered) |> sum()/nrow(instacart)) |> scales::percent()` of the orders, showing that only a small portion of instacart activity is captured by this data set. There are no first-time orders in this data set; this is on average order number `r instacart |> pull(order_number) |> mean() |> round(digits = 0)` for users (median order number `r instacart |> pull(order_number) |> median()` ). On average, users placed their orders `r instacart |> pull(days_since_prior_order) |> mean()` days after their previous one.

### Aisles & products
There are `r instacart |> pull(aisle_id) |> unique() |> length()` unique aisles included in the data set. The most ordered from aisles are the 

